---
output:
  word_document: default
  html_document: default
---
#Fields, Alex
##BAN 502
###Module 4 Assignment 3 - RF


####Task 3 - Q&A
The most important variable is *TotalDonations* with a score of 100.00 and *Mnths_Since_Last* at 98.25. The least important variable being *Total_Donated at 0*. 

####Task 5 - Q&A
When running the confustion matrix on the training set I was able to  see the model had an Accuracy of 0.8931, Sensitivity of 0.6080 and Specificity of 0.9825. 

####Task 6 - Q&A
When looking at the Naive model, it seems that the accuracy is a little better. This could be it is assuming all observations are in the majority. 

####Task 7 - Q&A
The testing data seems to be less accurate than the training set. It drops almost 10% while the p-value (significance) becomes larger which is not what you want to see. The Model for the Testing data is not what we would like to have. 

####Task 8 - Q&A
Comment on how this model might be used in the "real-world." Would you recommend this model
for real-world use? What if any concerns would you have about using the model?
**This model could be used for blood donations like we have tested but for accuracy we would need a lot more data. We coud also use this model in predicting the weather/temperature in a certain region. I would recommend this model, its not the best but it is still fairly accurate. My only concern is the need for more data when using this model.**


###Library
```{r message = FALSE}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(VIM) #visualizing missingness
library(ranger) #for random forests
library(caret)
library()

```



### Cleaning Data
```{r message = FALSE}
library(readr)
Blood <- read_csv("Blood.csv")

Blood = Blood %>% mutate(DonatedMarch = as.factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1" ))

```


###Training/Testing Split
```{r}
set.seed(1234)
train.rows = createDataPartition(y = Blood$DonatedMarch, p = 0.7, list =FALSE)
train = slice(Blood, train.rows)
test = slice(Blood, -train.rows)
```


###Random Forest Generation
```{r}
fit_control = trainControl(method = "cv",  
                           number = 10) #set up 10 fold cross-validation


set.seed(1233)  
rf_fit = train(DonatedMarch ~.,
                data = train, 
                method = "ranger", 
                importance = "permutation",
                trControl = fit_control,
                num.trees = 100)
```

###Validating variable importance
```{r}
varImp(rf_fit)
```


###Predictions on training
```{r}
predRF = predict.train(rf_fit, train)
head(predRF)
```

###Confusion Matrix on training
```{r}
confusionMatrix(predRF, train$DonatedMarch, positive = "Yes")
```


###Predictions on test
```{r}
predRF_test = predict(rf_fit, newdata = test)
```

###Confusion matrix on test
```{r}
confusionMatrix(predRF_test, test$DonatedMarch, positive = "Yes")
```




